---
title: "Homework 5, Diamonds Assignment (David)"
output: github_document
#########
# Script for Homework 5, Diamonds assignment, David
#########
---
##### Loading Libraries
```{r, message=FALSE}
# Loading libraries for the script, hide output
library(tidyverse)
```

##### 1) Print a data frame that contains the simple count of diamonds in each cut
```{r}
#Group the dataset and then count each cut
diamonds %>%
  group_by(cut) %>%
  count()
```

##### 2) Use dplyr functions to create a random subset (1% of total) of the diamonds dataset and check the results
```{r}
diamonds_subset <-sample_frac(diamonds,0.01)
head(diamonds_subset)
```

##### 3) Calculate the average size of the 100 largest diamonds in each clarity category (using the diamonds subset). 
##### Determine if there are 100 of each clarity category in the diamonds subset dataframe
```{r}
diamonds_subset %>%
  group_by(clarity) %>%
  count()
```
##### Since there are not 100 in each category, we will expand back to the original diamonds dataset
#### Determine if there are 100 in each clarity category in the original diamonds dataframe
```{r}
diamonds %>%
  group_by(clarity) %>%
  count()
```
##### We now have determined that we have to use the original dataframe to obtain the 100 largest diamonds in each clarity category. 
```{r}
#Grab the 100 largest diamonds in each category
largest_diamonds <- diamonds %>%
  arrange(desc(carat)) %>%
  group_by(clarity) %>%
  slice(1:100)
largest_diamonds %>%
  group_by(clarity) %>%
  count()
```
##### Calculate the average size of the largest diamond in each category
```{r}
largest_diamonds %>%
  group_by(clarity) %>%
  summarise(mean(carat))
```

##### 4) Make two scatter plots, width(x) vs length(y) and width(x) vs depth(z). The naming on the assignment suggests to plot width on the y-axis and length,depth on the x-axis. 
``` {r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = y, y = x)) +
  labs(title = "Diamond dimensions (in mm) width vs length", x = "length (y)", y = "width (x)")
```
``` {r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = z, y = x)) +
  labs(title = "Diamond dimensions (in mm) width vs depth", x = "depth (z)", y = "width (x)")
```

##### The plots show that there are a few outliers. We will filter the outliers out of the dataset.We will plot the results
```{r}
diamonds_filtered <-
  diamonds %>%
  filter(x > 3 & y < 20 & z>2 & z < 10)
head(diamonds_filtered)

##### Plot the filtered diamond results
```{r}
ggplot(data = diamonds_filtered) +
  geom_point(mapping = aes(x = y, y = x)) +
  labs(title = "Diamond dimensions (in mm) width vs length (filtered)", x = "length (y)", y = "width (x)")
```
``` {r}
ggplot(data = diamonds_filtered) +
  geom_point(mapping = aes(x = z, y = x)) +
  labs(title = "Diamond dimensions (in mm) width vs depth (filtered)", x = "depth (z)", y = "width (x)")
```




